---
title: 'Interaction: Approach x Large Use Cases'
author: "Julian Frattini"
date: "2025-08-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse) # R dialect for improved syntax and base functionality

library(brms) # Bayesian regression modeling
library(marginaleffects) # plotting of marginal and conditional effects

library(RColorBrewer) # color palettes for nominal data
path.figures <- "./../../figures/results/conditional/"
```

This notebook contains the analysis of the potential interaction effect between *TLR approach* (`approach`) and *Large Use Case* (`size`) on the outcome (here: `f2`).
The analysis provides insight into the hypothesis that some approaches perform differently well for different levels of use case size.

## Data Loading

First, load the data for the observational analysis.
The data set is procured by the [data-preparation.Rmd](./util/data-preparation.Rmd) notebook.
This notebook assembles and casts the data appropriately and stores the results in the `rq4tlr-data.csv` file.

```{r data-independent}
df <- read.csv('./../data/rq4tlr-data.csv')
```

Scale all numeric variables to make their coefficient distributions easier to sample from.

```{r data-scaling}
df <- df %>% 
  mutate(across(
    c("size", "outlinks", "maxinlinks", "maxsteplength", "sencomplexity"), 
    \(x) scale(x, center = TRUE))) %>% 
  mutate(
    Approach = factor(Approach, levels=c("FTLR", "LiSSA", "IR", "VSM", "LSI"), ordered = FALSE),
    Category = factor(Category, levels=c("Embedding", "LLM", "Traditional"), ordered = FALSE),
  )

str(df)
```

The output shows by what factors the numeric variables were centered and scaled.

## Analysis

The analysis of the data employs a regression model in a Bayesian framework.
The utilized formula only assumes an intercept (i.e., a baseline), an offset per dataset, and the interaction effect between `approach` and the current variable of interest.

```{r formula}
f <- F2 ~ 0 + Intercept + Dataset + Approach*size
get_prior(f, data=df, family=zero_one_inflated_beta)
```

Given this formula, select uninformative distributions for each of the coefficient distributions.

```{r priors}
p1 <- c(
  prior(normal(0, 1), class = b, coef = "Intercept"),
  prior(normal(0, 0.5), class = b),
  prior(gamma(4, 1), class = phi),
  prior(beta(7, 2), class = coi), 
  prior(beta(5, 2), class = zoi) 
)
```

To confirm the eligibility of the prior distributions, perform a prior predictive check.

```{r model-sample-priors}
m.prior <-
  brm(data = df, family = zero_one_inflated_beta, f, prior = p1,
    iter = 4000, warmup = 1000, chains = 4, cores = 4,
    seed = 4, sample_prior="only",
    file = "../fits/int.f2.approach.size.prior"
  )

brms::pp_check(m.prior, ndraws=100)
```

Due to the number of interaction effects and the relatively wide priors, the estimation ($y_{rep}$) clusters around the extremes of the spectrum.
This is, however, expected and will be overwhelmed by the number of data points.

Given the eligibility of the priors, train the full model by updating the coefficient distributions in light of the observed data.

```{r model}
m <-
  brm(data = df, family = zero_one_inflated_beta, f, prior = p1,
    iter = 4000, warmup = 1000, chains = 4, cores = 4, seed = 4,
    file = "../fits/int.f2.approach.size"
  )

brms::pp_check(m, ndraws=100)
```

The posterior predictive check confirms that the model trained properly.

### Evaluation

To evaluate the trained model and draw conclusions about the hypothesized conditional effect, we first look at the updated path coefficients in the model summary.
These coefficients are in relation to one selected baseline approach (here: `FTLR`) and represent the average slope difference.

```{r path-coefficients}
summary(m)
```

More contextualized insights can be derived by plotting the conditional effect.

```{r conditional-effect}
c.eff <- conditional_effects(m, effects="size:Approach", 
                    prob = 0.5, method="posterior_epred")

approaches = c("VSM", "LSI", "FTLR", "LiSSA", "LiSSA_IR-only")

c.eff$`size:Approach` %>% 
  mutate(Approach = as.character(Approach)) %>% 
  mutate(Approach = ifelse(Approach == "IR", "LiSSA_IR-only", Approach)) %>% 
  mutate(Approach = factor(Approach, levels = approaches, ordered = TRUE)) %>% 
  mutate(size = size*11.3 + 10.9) %>% 
  ggplot(aes(x = size)) +
    geom_line(aes(y = estimate__, color = Approach)) +
    geom_ribbon(aes(ymin = lower__, ymax = upper__, fill = Approach), alpha = 0.2) +
    scale_color_manual(values = brewer.pal(n = 5, name = "Set1"), limits = approaches) +
    scale_fill_manual(values = brewer.pal(n = 5, name = "Set1"), limits = approaches) +
    labs(x = "Total number of steps", y = "Average F2 Value") +
    ylim(0.2, 0.5) +
    theme_bw() +
    theme(legend.position = "bottom")

ggsave(
 file = file.path(path.figures, "obs-approach-size.pdf"), device = "pdf", 
 width = 5.3, height = 3.5)
```

The results invite the following interpretation:

- LiSSA can cope with large use cases better, as GPT-4o is the most capable approach in interpreting semantics.LiSSA seems to be able to retrieve the different referred steps anyways (probably because it is the only classification-based approach: it classifies each combination of source and target retrieved by the Lissa_IR independently, instead of retrieving everything at once as VSM LSI and IR).
- FTLR works on single steps and aggregates from the code side and is, thus, less sensitive than the other embedding-based approaches (LSI,IR) to many different concerns in the source artifact.
- More words seem to help VSM, maybe because this way TF-IDF works better.

```{r export-figure-data}
conditional_effects(m, 
                    effects="size:Approach",
                    method="posterior_epred")$`size:Approach` %>% 
  select(c("size", "Approach", "estimate__", "lower__", "upper__")) %>% 
  rename(qf = "size") %>% 
  write_csv(file = "../data/estimates/conditional-size.csv")
```
