---
title: 'Interaction: Approach x UI Design Details'
author: "Julian Frattini"
date: "2025-08-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse) # R dialect for improved syntax and base functionality

library(brms) # Bayesian regression modeling
library(marginaleffects) # plotting of marginal and conditional effects
```

This notebook contains the analysis of the potential interaction effect between *TLR approach* (`approach`) and *Contains UI Design Details* (`uidesign`) on the outcome (here: `f2`).
The analysis provides insight into the hypothesis that some approaches perform differently well for different degrees of UI details in the requirements.

## Data Loading

First, load the data for the observational analysis.
The data set is procured by the [data-preparation.Rmd](./util/data-preparation.Rmd) notebook.
This notebook assembles and casts the data appropriately and stores the results in the `rq4tlr-data.csv` file.

```{r data-independent}
df <- read.csv('./../data/rq4tlr-data.csv')
```

Scale all numeric variables to make their coefficient distributions easier to sample from.

```{r data-scaling}
df <- df %>% 
  mutate(across(
    c("size", "outlinks", "maxinlinks", "maxsteplength", "sencomplexity"), 
    \(x) scale(x, center = TRUE))) %>% 
  mutate(
    Approach = factor(Approach, levels=c("FTLR", "LiSSA", "IR", "VSM", "LSI"), ordered = FALSE),
    Category = factor(Category, levels=c("Embedding", "LLM", "Traditional"), ordered = FALSE),
  )

str(df)
```

The output shows by what factors the numeric variables were centered and scaled.

## Analysis

The analysis of the data employs a regression model in a Bayesian framework.
The utilized formula only assumes an intercept (i.e., a baseline), an offset per dataset, and the interaction effect between `approach` and the current variable of interest.

```{r formula}
f <- F2 ~ 0 + Intercept + Dataset + Approach*uidesign
get_prior(f, data=df, family=zero_one_inflated_beta)
```

Given this formula, select uninformative distributions for each of the coefficient distributions.

```{r priors}
p1 <- c(
  prior(normal(0, 1), class = b, coef = "Intercept"),
  prior(normal(0, 0.5), class = b),
  prior(gamma(4, 1), class = phi),
  prior(beta(7, 2), class = coi), 
  prior(beta(5, 2), class = zoi) 
)
```

To confirm the eligibility of the prior distributions, perform a prior predictive check.

```{r model-sample-priors}
m.prior <-
  brm(data = df, family = zero_one_inflated_beta, f, prior = p1,
    iter = 4000, warmup = 1000, chains = 4, cores = 4,
    seed = 4, sample_prior="only",
    file = "../fits/int.f2.approach.uidesign.prior"
  )

brms::pp_check(m.prior, ndraws=100)
```

Due to the number of interaction effects and the relatively wide priors, the estimation ($y_{rep}$) clusters around the extremes of the spectrum.
This is, however, expected and will be overwhelmed by the number of data points.

Given the eligibility of the priors, train the full model by updating the coefficient distributions in light of the observed data.

```{r model}
m <-
  brm(data = df, family = zero_one_inflated_beta, f, prior = p1,
    iter = 4000, warmup = 1000, chains = 4, cores = 4, seed = 4,
    file = "../fits/int.f2.approach.uidesign"
  )

brms::pp_check(m, ndraws=100)
```

The posterior predictive check confirms that the model trained properly.

### Evaluation

To evaluate the trained model and draw conclusions about the hypothesized conditional effect, we first look at the updated path coefficients in the model summary.
These coefficients are in relation to one selected baseline approach (here: `FTLR`) and represent the average slope difference.

```{r path-coefficients}
summary(m)
```

More contextualized insights can be derived by plotting the conditional effect.

```{r conditional-effect}
conditional_effects(m, effects="uidesign:Approach", 
                    prob = 0.5, method="posterior_epred")
```
