---
title: "Observational Analysis"
author: "Anonymous"
date: '2025-02-17'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse) # R dialect for improved syntax and base functionality

library(brms) # Bayesian regression modeling
library(marginaleffects) # plotting of marginal and conditional effects
```

This notebook contains the observational analysis of the impact of requirements quality (RQ) on trace link recovery (TLR).

## Data Loading

First, load the data for the observational analysis.
The data set is procured by the [data-preparation.Rmd](./util/data-preparation.Rmd) notebook.
This notebook assembles and casts the data appropriately and stores the results in the `rq4tlr-dtata.csv` file.

```{r data-independent}
df <- read.csv('./../data/rq4tlr-data.csv')
```

Scale all numeric variables to make their coefficient distributions easier to sample from.

```{r data-scaling}
df <- df %>% 
  mutate(across(c("size", "outlinks", "maxinlinks", "maxsteplength", "sencomplexity"), scale, center = TRUE))

str(df)
```

## Analysis

The analysis of the data employs a regression model in a Bayesian framework.

### Distribution

The three outcome variables `Precision`, `Recall`, and `F1` are all bounded $\in [0, 1]$.
This makes a **Beta** distribution an eligible target distribution of the response variables.
To determine which variant of the Beta distribution family can be chosen, we visualize the raw distribution.

```{r outcome-distribution}
df %>% 
  select(Precision) %>% 
  ggplot(aes(x = Precision)) +
    geom_histogram(bins = 30)
```

A base Beta distribution would require that all values are bounded within $]0; 1[$.
The observed values for the three outcome variables show a clear zero- and one-inflation.
Hence, we chose the `zero_one_inflated_beta` distribution from the `brms` library.

### Formula

Next, we define regression formulae that put the outcome variable of interest in relation to the predictors with which we assume a causal relationship.

```{r formula}
f <- Precision ~ 0 + Intercept + Dataset + sencomplexity + maxsteplength + size + inconsistentloa + uidesign + clarification + alternative + coherence + happy + lostfr + anaphora + nonnounstart + passive + negation + outlinks + maxinlinks
```

The `0 + Intercept` term instead of `1 +` or even complete omission stems from the fact that the regression formula contains at least one predictor that is not centered around 0.

### Priors

The selected regression formula requires the definition of prior distributions for the following parameters.

```{r priors-eligible}
get_prior(f, data=df, family=zero_one_inflated_beta)
```

We select the following priors:

- $N(0, 1)$ for the intercept, to allow a flexible baseline
- $N(0, 0.5)$ centered, uninformative priors for all factors
- $Gam(4, 1)$ for the concentration parameter $\phi$
- $Beta(7, 2)$ for the zero-inflation
- $Beta(4, 2)$ for the one-inflation

```{r priors1}
p1 <- c(
  prior(normal(0, 1), class = b, coef = "Intercept"),
  prior(normal(0, 0.5), class = b),
  prior(gamma(4, 1), class = phi),
  prior(beta(7, 2), class = coi), 
  prior(beta(5, 2), class = zoi) 
)
```

### Prior Predictive Check

The prior predictive check helps determine the eligibility of the priors.

```{r model-sample-priors}
m1.prior <-
  brm(data = df, family = zero_one_inflated_beta, f, prior = p1,
    iter = 4000, warmup = 1000, chains = 4, cores = 4,
    seed = 4, sample_prior="only",
    file = "../fits/obs.precision.prior"
  )
```

In our particular case, the large number of predictors will, however, push the accumulated value of $\mu$ in the Beta-distribution to the to extremes.

```{r prior-predictive-check}
brms::pp_check(m1.prior, ndraws=100)
```

The figure shows extreme concentrations of expected values $y_{rep}$ on the borders of the output space (at 0 and 1).
Instead, we put more emphasis on the posterior predictive check.

### Training

Assuming the priors are reliable, we train the model by updating the coefficient parameters with the data `df`.

```{r model}
m1 <-
  brm(data = df, family = zero_one_inflated_beta, f, prior = p1,
    iter = 4000, warmup = 1000, chains = 4, cores = 4, seed = 4,
    file = "../fits/obs.precision"
  )
```

To check whether the training completed successfully, we inspect the posterior predictive check.

```{r posterior-predictive-check}
brms::pp_check(m1, ndraws=100)
```

The output confirms that the model parameters have been updated properly and that the predicted outcomes $y_{rep}$ are close the the actual outcome $y$.

### Evaluation

Finally, we evaluate the model.

#### Coefficients

First, we look at the updated path coefficients in the model summary.

```{r path-coefficients}
summary(m1)
```

These coefficients can also be visualized via a `mcmc_plot`.

```{r mcmc-plot}
mcmc_plot(m1, variable=c("b_Datasetitrust", "b_sencomplexity", "b_maxsteplength", "b_size", "b_inconsistentloa", "b_uidesign", "b_clarification", "b_alternative", "b_coherence", "b_happyTRUE", "b_lostfrTRUE", "b_anaphora", "b_nonnounstart", "b_passive", "b_negation", "b_outlinks", "b_maxinlinks"))
```

Variables, where the 50% confidence interval of the intercept does not overlap 0, weakly impact the response variable.
Variables, where the 95% confidence interval of the intercept does not overlap 0, strongly impact the response variable.

#### Marginal Distributions

For a clearer picture, we can plot the marginal distributions of the outcome variable when manipulating one specific factor while keeping the others at a constant level.

```{r marginal-inconsistenloa}
conditional_effects(m1, effects="inconsistentloa", method="posterior_epred")
```

```{r marginal-nonnounstart}
conditional_effects(m1, effects="nonnounstart", method="posterior_epred")
```

```{r marginal-outlinks}
conditional_effects(m1, effects="outlinks", method="posterior_epred")
```

## Store Data

Finally, store the estimates of the posterior distributions.

```{r export-estimates}
summary(m1)$fixed %>% 
  select(Estimate, Est.Error) %>% 
  rename(all_of(c(
    mean = "Estimate", 
    var = "Est.Error"))) %>% 
  rownames_to_column("coef") %>% 
  write_csv(file = "../data/estimates/obs-precision.csv")
```

