---
title: "Experimental Analysis"
author: "Anonymous"
date: '2025-02-26'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

library(brms)
library(marginaleffects)

library(patchwork)
```

In this Notebook, we conduct an experimental analysis.
As opposed to the observational analysis, the experimental analysis allows to fully control a factor and, hence, isolate its effect on the response variable of interest.

```{r load-data}
d <- read.csv('./../data/rq4tlr-data.csv')

head(df)
```


## Factor Selection

We can only conduct an experimental analysis of a specific factor if

1. the factor is sufficiently wide distributed in the data, and
2. we can reliably manipulate the factor without requiring domain knowledge.

Only one factor in our data meets both criteria: the ratio of sentences representing clarifications (`clarification`).
The factor is at least spread a little in the `iTrust` data set.

```{r visualize-clarifications}
d %>% 
  filter(Dataset == "itrust") %>% 
  ggplot(aes(x = clarification, fill = Dataset)) +
  geom_histogram(stat="count")
```

```{r}
d.treatment <- read.csv('../../data/TLR_results/FTLR/itrustClarificationCleaned_OPT_m52f42_FTLRMcCd_perArtifactMetrics.csv', sep=';') %>%
  mutate(Dataset = 'itrust') %>% 
  rowwise() %>% 
  mutate(File = strsplit(sourceID, '\\.')[[1]][1]) %>% 
  select(!c(sourceID, tp, fp, fn, LAG))

d.treatment$clarification = 0.0

head(d.treatment)
```

```{r}
d.control <- d %>% 
  filter(Dataset == "itrust") %>% 
  select(c(Dataset, File, clarification, Precision, Recall, F1, AP))

head(d.control)
```

```{r merge-data}
d <- rbind(d.control, d.treatment)
```

```{r calculate-f2}
calc_f_beta <- function(beta, precision, recall) {
  f <- (1+beta^2) * ((precision * recall)/(beta^2*precision + recall))
  return(ifelse(is.nan(f), 0, f))
}

d <- d %>% 
  mutate(F2 = calc_f_beta(2, Precision, Recall))
```


## Data Analysis

```{r outcome-distribution}
d %>% 
  pivot_longer(cols=c(Precision, Recall, F1, F2, AP), names_to = "Metric") %>% 
  ggplot(aes(x = value, fill=Metric)) +
    geom_histogram()
```

```{r formula}
f <- bf(mvbind(Precision, Recall, F1, F2, AP) ~ 0 + Intercept + clarification + (1|File)) + set_rescor(FALSE)
#f <- bf(mvbind(Precision, Recall) ~ 0 + Intercept + clarification + (1|File)) + set_rescor(FALSE)
```

```{r show-priors}
get_prior(f, d, family=zero_one_inflated_beta)
```

```{r select-priors}
p1 <- c(
  prior(normal(0, 0.3), class = b, resp=Precision),
  prior(normal(0, 0.5), class = b, resp=Recall),
  prior(normal(0, 0.5), class = b, resp=F1),
  prior(normal(0, 0.5), class = b, resp=F2),
  prior(normal(0, 0.5), class = b, resp=AP),
  
  # phi parameter
  prior(gamma(4, 1), class = phi, resp=Precision),
  prior(gamma(4, 1), class = phi, resp=Recall),
  prior(gamma(4, 1), class = phi, resp=F1),
  prior(gamma(4, 1), class = phi, resp=F2),
  prior(gamma(4, 1), class = phi, resp=AP),
  
  # coi parameter
  prior(beta(5, 2), class = coi, resp=Precision), 
  prior(beta(5, 2), class = coi, resp=Recall), 
  prior(beta(5, 2), class = coi, resp=F1), 
  prior(beta(5, 2), class = coi, resp=F2), 
  prior(beta(5, 2), class = coi, resp=AP), 
  
  # zoi parameter
  prior(beta(3, 2), class = zoi, resp=Precision),
  prior(beta(3, 2), class = zoi, resp=Recall),
  prior(beta(3, 2), class = zoi, resp=F1),
  prior(beta(3, 2), class = zoi, resp=F2),
  prior(beta(3, 2), class = zoi, resp=AP),
  
  # standard deviation parameter
  prior(weibull(2, 0.5), class = sd, resp=Precision),
  prior(weibull(2, 0.5), class = sd, resp=Recall),
  prior(weibull(2, 0.5), class = sd, resp=F1),
  prior(weibull(2, 0.5), class = sd, resp=F2),
  prior(weibull(2, 0.5), class = sd, resp=AP)
)
```

```{r model-sample-priors}
m1.prior <-
  brm(data = d, family = zero_one_inflated_beta, f, prior = p1,
    iter = 4000, warmup = 1000, chains = 4, cores = 4,
    seed = 4, sample_prior="only",
    file = "../fits/exp.prior"
  )
```

```{r prior-predictive-checks}
prior.precision <- brms::pp_check(m1.prior, ndraws=100, resp="Precision")
prior.recall <- brms::pp_check(m1.prior, ndraws=100, resp="Recall")
prior.f1 <- brms::pp_check(m1.prior, ndraws=100, resp="F1")
prior.f2 <- brms::pp_check(m1.prior, ndraws=100, resp="F2")
prior.ap <- brms::pp_check(m1.prior, ndraws=100, resp="AP")

(prior.precision | prior.recall) / (prior.f1 | prior.f2) / prior.ap
```

```{r model}
m1 <-
  brm(data = d, family = zero_one_inflated_beta, f, prior = p1,
    iter = 4000, warmup = 1000, chains = 4, cores = 4, seed = 4,
    file = "../fits/exp"
  )
```

```{r posterior-predictive-check}
posterior.precision <- brms::pp_check(m1, ndraws=100, resp="Precision")
posterior.recall <- brms::pp_check(m1, ndraws=100, resp="Recall")
posterior.f1 <- brms::pp_check(m1, ndraws=100, resp="F1")
posterior.f2 <- brms::pp_check(m1, ndraws=100, resp="F2")
posterior.ap <- brms::pp_check(m1, ndraws=100, resp="AP")

(posterior.precision | posterior.recall) / (posterior.f1 | posterior.f2) / posterior.ap
```

```{r path-coefficients}
summary(m1)
```

```{r mcmc-plot}
mcmc_plot(m1, variable=c("b_Precision_clarification", "b_Recall_clarification", "b_F1_clarification", "b_F2_clarification", "b_AP_clarification"))
```

```{r}
dodge = 0.8 # distance between errorbars of each metric

ci1.width = 0.6 # size of the whiskers of the outer confidence interval
ci2.width = 0.4 # size of the whiskers of the inner confidence interval

ci1.thic = 0.1 # thickness of the outer confidence interval
ci2.thic = 1.3 # thickness of the inner confidence interval

summary(m1)$fixed %>% 
  select(Estimate, Est.Error) %>% 
  rownames_to_column("coef") %>% 
  filter(endsWith(coef, "clarification")) %>% 
  rowwise() %>% 
  mutate(coef = strsplit(coef, "_")[[1]][1]) %>% 
  rename(all_of(c(
    avg = "Estimate",
    var = "Est.Error"
  ))) %>% 
  ggplot(aes(y = "a", color = coef)) +
    geom_point(aes(x = avg), position=position_dodge(width=dodge), size = 3) +
    geom_errorbar(aes(xmin = avg-2*var, xmax = avg+2*var), 
                  width=ci1.width, linewidth=ci1.thic, 
                  position=position_dodge(width=dodge)) +
    geom_errorbar(aes(xmin = avg-var, xmax = avg+var), 
                  width=ci2.width, linewidth=ci2.thic, 
                  position=position_dodge(width=dodge)) + 
    geom_vline(xintercept=0, color = "grey") +
    theme(axis.title.y=element_blank(), axis.ticks.y = element_blank(), axis.text.y = element_blank()) +
    labs(
      x = "Posterior distribution of the coefficient of the Clarification factor",
      color = "Metric")
```


### Marginal Effects

```{r marginal}
conditional_effects(m1, effects="clarification", method="posterior_epred")
```
