---
title: "Multiple Conditional Effects"
author: "Julian Frattini"
date: "2025-08-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(RColorBrewer)

path.data <- "./../data/estimates/"
path.figures <- "./../../figures/results/conditional/"
```

This notebook visualizes how the TLR approach moderates the three effect of the three quality factors *Inconsistent Levels of Abstraction* (`inconsistentloa`), *Tangled Requirements* (`outlinks`), and *Large Use Case* (`size`) on the $F_2$-score.

## Data Preparation

First, load the respective data, annotate the quality factor as `qFactor`, and combine the three data sets.

```{r load-data}
d.inconsistentloa <- read.csv(file.path(path.data, "conditional-inconsistentloa.csv")) %>% 
  mutate(qFactor = "inconsistentloa")
d.outlinks <- read.csv(file.path(path.data, "conditional-outlinks.csv")) %>% 
  mutate(qFactor = "outlinks")
d.size <- read.csv(file.path(path.data, "conditional-size.csv")) %>% 
  mutate(qFactor = "size")

d <- rbind(d.inconsistentloa, d.outlinks, d.size) %>% 
  mutate(
    Approach = factor(Approach, levels=c("FTLR", "LiSSA", "IR", "VSM", "LSI"), 
                      labels = c("FTLR", "LiSSA", "LiSSA_IR-only", "VSM", "LSI"), 
                      ordered = FALSE),
    qf = ifelse(qFactor == "outlinks", 
                qf*2.46 + 3.15,
                ifelse(qFactor == "size", 
                       qf*11.3 + 10.9,
                       qf)),
    qFactor = factor(qFactor, levels = c("inconsistentloa", "outlinks", "size"),
                     labels = c("Fraction of Steps on an inconsistent Level of Abstraction", 
                                "Number of outgoing Traceability Links",
                                "Total Number of Steps"), ordered = FALSE)
    )
```

## Data Visualization

Visualize the conditional effects in a line plot with ribbons.

```{r visualize-plot}
approaches = c("VSM", "LSI", "FTLR", "LiSSA", "LiSSA_IR-only")

ggplot(d, aes(x = qf)) +
  geom_line(aes(y = estimate__, color = Approach)) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__, fill = Approach), alpha = 0.1) +
  facet_wrap(~qFactor, scale = "free_x") +
  labs(y = "Average F2 Value") +
  scale_color_manual(values = brewer.pal(n = 5, name = "Set1"), limits = approaches) +
  scale_fill_manual(values = brewer.pal(n = 5, name = "Set1"), limits = approaches) +
  theme_bw() +
  theme(axis.title.x = element_blank())
```

```{r export-figure}
ggsave(
  file = file.path(path.figures, "obs-conditional-inconsistentloa-outlinks-size.pdf"), 
  device = cairo_pdf, width = 12, height = 3)
```
