---
title: "Causal DAG"
author: "Anonymous"
date: '2025-02-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggdag)
```

This notebook visualizes the causal assumptions about the collected data in the RQ4TLR study.

```{r variables-structure}
vars = data.frame(varname = character(), # short variable name (e.g., perf)
                  fullname = character(), # full, expressive name of the variable (e.g., TLR performance)
                  x = double(), y = double(), # x and y coordinates of this variable in the figure
                  stringsAsFactors=FALSE)

```

Add each variable to consider in the analysis to the `vars` data frame.

```{r variables}
vars <- vars %>% add_row(varname = "perf", fullname = "TLR Performance", x = 2, y = 0)

vars <- vars %>% add_row(varname = "dataset", fullname = "Data Set", x = -1, y = 0)
vars <- vars %>% add_row(varname = "sencomplexity", fullname = "Maximum Step Complexity", x = 0, y = 0.5)
vars <- vars %>% add_row(varname = "maxsteplength", fullname = "Maximum Step Length", x = 0, y = 0.3)
vars <- vars %>% add_row(varname = "size", fullname = "Use Case Size", x = 0, y = 0.1)
vars <- vars %>% add_row(varname = "inconsistentloa", fullname = "Inconsistent Level of Abstraction", x = 0, y = -0.1)
vars <- vars %>% add_row(varname = "uidesign", fullname = "Contains UI Design Details", x = 0, y = -0.2)
vars <- vars %>% add_row(varname = "clarification", fullname = "contains clarification", x = 0, y = -0.3)
vars <- vars %>% add_row(varname = "alternative", fullname = "Contains Alternative", x = 0, y = -0.4)
vars <- vars %>% add_row(varname = "coherence", fullname = "Level of Coherence", x = 0, y = -0.5)


vars <- vars %>% add_row(varname = "happy", fullname = "Happy Use Case", x = 0, y = -0.6)
vars <- vars %>% add_row(varname = "lostfr", fullname = "Mislocated Functional Requirements", x = 0, y = -0.7)
vars <- vars %>% add_row(varname = "anaphora", fullname = "Anaphora", x = 0, y = -0.8)
vars <- vars %>% add_row(varname = "nonnounstart", fullname = "Non-noun Start", x = 0.4, y = -0.9)
vars <- vars %>% add_row(varname = "passive", fullname = "Passive Voice", x = 0, y = -1)
vars <- vars %>% add_row(varname = "negation", fullname = "Negation", x = 0, y = -1.1)
vars <- vars %>% add_row(varname = "outlinks", fullname = "Scatter", x = 0, y = -1.2)
vars <- vars %>% add_row(varname = "maxinlinks", fullname = "Tangle", x = 0, y = -1.3)

```

Extract named lists out of the data frame (one for the full name, one for the x and one for the y positions).

```{r variables-data}
map.name <- setNames(vars$fullname, vars$varname)
map.x <- setNames(vars$x, vars$varname)
map.y <- setNames(vars$y, vars$varname)
```

Visualize the causal directed, acyclic graph from the variables.

```{r dag}
dag <- dagify(
  perf  ~ dataset + sencomplexity + maxsteplength + size + inconsistentloa + uidesign + clarification + alternative + coherence + happy + lostfr + anaphora + nonnounstart + passive + negation + outlinks + maxinlinks,
  
  size ~ dataset,
  maxsteplength ~ dataset,
  sencomplexity ~ dataset + maxsteplength,
  inconsistentloa ~ dataset,
  uidesign ~ dataset,
  clarification ~ dataset,
  alternative ~ dataset,
  coherence ~ dataset,
  happy ~ dataset,
  lostfr ~ dataset,
  anaphora ~ dataset,
  nonnounstart ~ dataset + passive,
  passive ~ dataset,
  negation ~ dataset,
  outlinks ~ dataset,
  maxinlinks ~ dataset,
  
  outcome = "perf", # determine TLR performance as the response variable
  labels = map.name,
  coords = list(x = map.x, y = map.y)
)

ggdag_status(dag, use_labels = "label", text = FALSE) +
  guides(fill = "none", color = "none") + 
  theme_dag()
```

The following table justifies the inclusion of each relationship between two variables in the causal DAG:

| Relationship | Reasoning |
|---|---|
| `dataset` $\rightarrow$ `TLR performance` | The TLR algorithm may perform differently on different data sets due to their differing complexity, etc. |
| `maximum step complexity` $\rightarrow$ `TLR performance` | The more complex the steps of a use case are, the more difficult it is to trace the use case to source code elements. |
| `maximum step length` $\rightarrow$ `TLR performance` | The longer the steps of a use case are, the more difficult it is to trace the use case to source code elements. |
| `number of steps` $\rightarrow$ `TLR performance` | The more steps a use case has, the more difficult it is to trace the use case to all source code elements. |

The following variables were ignored due to their imbalanced distribution in the data set:

- *Meaningless Actors* (`meanuc`): the variable perfectly correlates with the `dataset` variable and, therefore, bears no information
- *Functional Duplication* (`duplicate`): the data set did not contain a single instance of functional duplication
- *Meaningless Use Case* (`meanuc`): the data set contained too few instances of meaningless use cases
- *Use Case Naming Problems* (`nameproblem`): there are only a few instances in the iTrust data set with naming problems
- *Inappropriate Scope* (`scope`): there are hardly any instances of scoping problems encountered in the data set
- *Incoherent Test Order* (`textorder`): there are only a few instances in the iTrust data set with an incoherent text order
- *Inputs or Outputs not quantified* (`iounquantified`): there are only a few instances in the etour data set where the inputs/outputs are not quantified
- *Contains Actor-Actor Interaction* (`actoractor`): there are hardly any instances of actor-actor interactions encountered in the data set
- *Contains justifications* (`justification`): the data set did not contain a single justification
- *Coordination Ambiguity* (`coordambiguity`): there are only very few instances of coordination ambiguity at all
- *Optional* (`optional`): there are too few instances of optional parts in the data set
- *Incomplete Comparisons*: there are too few instances of incomplete comparisons and they only occur in the eTour data set

