---
title: "Causal DAG"
author: "Anonymous"
date: '2025-02-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggdag)
```

This notebook visualizes the causal assumptions about the collected data in the RQ4TLR study.

```{r variables-structure}
vars = data.frame(varname = character(), # short variable name (e.g., perf)
                  fullname = character(), # full, expressive name of the variable (e.g., TLR performance)
                  x = double(), y = double(), # x and y coordinates of this variable in the figure
                  stringsAsFactors=FALSE)

#```

#Add each variable to consider in the analysis to the `vars` data frame.

#```{r variables}
vars <- vars %>% add_row(varname = "perf", fullname = "TLR performance", x = 2, y = 0)

vars <- vars %>% add_row(varname = "dataset", fullname = "data set", x = -1, y = 0)
vars <- vars %>% add_row(varname = "maxcomplex", fullname = "maximum step complexity", x = 0, y = 0.5)
vars <- vars %>% add_row(varname = "maxlength", fullname = "maximum step length", x = 0, y = 0.3)
vars <- vars %>% add_row(varname = "steps", fullname = "number of steps", x = 0, y = 0.1)
vars <- vars %>% add_row(varname = "ilos", fullname = "inconsistent level of abstraction", x = 0, y = -0.1)
vars <- vars %>% add_row(varname = "cuid", fullname = "contains ui design details", x = 0, y = -0.2)
vars <- vars %>% add_row(varname = "clarify", fullname = "contains clarifications", x = 0, y = -0.3)
vars <- vars %>% add_row(varname = "alt", fullname = "contains alternatives", x = 0, y = -0.4)
vars <- vars %>% add_row(varname = "coherence", fullname = "coherence", x = 0, y = -0.5)

#```

#Extract named lists out of the data frame (one for the full name, one for the x and one for the y positions).

#```{r variables-data}
map.name <- setNames(vars$fullname, vars$varname)
map.x <- setNames(vars$x, vars$varname)
map.y <- setNames(vars$y, vars$varname)
#```

#Visualize the causal directed, acyclic graph from the variables.

#```{r dag}
dag <- dagify(
  perf  ~ dataset + maxcomplex + maxlength + steps + ilos + cuid + clarify + alt + coherence,
  
  steps ~ dataset,
  maxlength ~ dataset,
  maxcomplex ~ dataset + maxlength,
  ilos ~ dataset,
  cuid ~ dataset,
  clarify ~ dataset,
  alt ~ dataset,
  coherence ~ dataset,
  
  outcome = "perf", # determine TLR performance as the response variable
  labels = map.name,
  coords = list(x = map.x, y = map.y)
)

ggdag_status(dag, use_labels = "label", text = FALSE) +
  guides(fill = "none", color = "none") + 
  theme_dag()
```

The following table justifies the inclusion of each relationship between two variables in the causal DAG:

| Relationship | Reasoning |
|---|---|
| `dataset` $\rightarrow$ `TLR performance` | The TLR algorithm may perform differently on different data sets due to their differing complexity, etc. |
| `maximum step complexity` $\rightarrow$ `TLR performance` | The more complex the steps of a use case are, the more difficult it is to trace the use case to source code elements. |
| `maximum step length` $\rightarrow$ `TLR performance` | The longer the steps of a use case are, the more difficult it is to trace the use case to source code elements. |
| `number of steps` $\rightarrow$ `TLR performance` | The more steps a use case has, the more difficult it is to trace the use case to all source code elements. |

The following variables were ignored due to their imbalanced distribution in the dataset:

- `meaningless.actor`: the variable perfectly correlates with the `dataset` variable
- `coordination.ambiguity`: there are too few instances of values beyond 0.0.
- `happy.ucs`
- `meaningless.ucs`
- `contains.optional`
- `incomplete comparisons`
- `functional.duplication`
- `user.case.naming.problem` -> candidate for manipulation
- `inappropriate.scope`
- `incoherent.text.order`
- `inputs.or.outputs.not.quantified`
- `actor.actor.interaction`
- `contains.justifications`

Candidates for consideration:

- `tangled` (once transformed to number)
- `scattered` (once transformed to number)
- `nfr`?
- `contains.anaphora`
- `starts.with.noun` / `passive`
- `negation`

