---
title: "Data Preparation"
author: "Anonymous"
date: '2025-02-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

# location of the TLR results
data.path <- '../../data/TLR_results/'
```

This notebook prepares the observational data for its subsequent analysis.

## Indepedent Data

Firstly, load the `csv` sheet containing all observations, which contains both manually and automatically annotated variables.

```{r data-loading}
df.ind <- read.csv('../../data/output/rq4tlr-merged.csv')
```

### Datatype Correction

Several variables are Boolean but parsed as characters (`"TRUE"` and `"FALSE"`) in the data sheet.
To mitigate this, cast these variables to actual Boolean values.

```{r data-casting-bool}
vars.bool <- c("duplicate", "nameproblem", "scope", "textorder", "iounquantified", "actoractor", "justification", "happy", "meanuc", "lostfr")

to_bool <- function(value) {
  return(value == "True")
}

df.ind <- df.ind %>% 
  mutate_at(vars.bool, to_bool)
```

The resulting independent data contains all variables in their correct format.

```{r data-status}
str(df.ind)
```

### Value Correction

The values in the `coherence` column contain the value -1.0 whenever the use case subflow contained no more than 1 step.
In this case, the coherence value is not meaningful because there cannot be coherence of any sentence to any predecessor.

```{r coherence-hist}
df.ind %>% ggplot(aes(x = coherence)) +
  geom_histogram()
```

The value -1.0 would be misinterpreted by a regression model.
Replacing those values with `NA` will discard them for the regression analysis.
Instead, we replace those values with 1: a subflow with only one step has perfect coherence by default.

```{r coherence-adoption}
df.ind <- df.ind %>% 
  mutate(coherence = ifelse(coherence == -1, 1, coherence))
```

## Dependent Data

The dependent variables are located in the `./data/TLR_results/` directory.
The trace link recovery results per data set are stored in separate files. 

```{r load-tlr-function}
load.tlr.results <- function(filename.etour, filename.itrust, approach, category, separator=';') {
  df.dep.etour <- read.csv(file.path(data.path, filename.etour), sep=separator) %>%
    mutate(Dataset = 'etour')
  df.dep.itrust <- read.csv(file.path(data.path, filename.itrust), sep=separator) %>% 
    mutate(Dataset = 'itrust')
  
  # join the TLR results from the two data sets and declare the approach and category
  df.dep <- rbind(df.dep.etour, df.dep.itrust) %>% 
    mutate(
      Approach = approach,
      Category = category
    )
  
  # clean the sourceID column
  df.dep <- df.dep %>% 
    rowwise() %>% 
    mutate(File = strsplit(sourceID, '\\.')[[1]][1]) %>%
    select(!(sourceID))
  
  return(df.dep)
}
```

Load the data for each of the approaches.

```{r load-tlr}
tlr.ftlr <- load.tlr.results(filename.etour = 'FTLR/etour_OPT_m57f41_FTLRMcCd_perArtifactMetrics.csv',
                 filename.itrust = 'FTLR/itrust_OPT_m52f42_FTLRMcCd_perArtifactMetrics.csv',
                 approach = 'FTLR', category = 'Embedding')

tlr.ir <- load.tlr.results(filename.etour = 'LiSSA/eTour_LiSSA_IR_Only_perArtifactMetrics.csv',
                 filename.itrust = 'LiSSA/iTrust_LiSSA_IR_Only_perArtifactMetrics.csv',
                 approach = 'IR', category = 'Embedding', separator = ',')

tlr.lissa <- load.tlr.results(filename.etour = 'LiSSA/eTour_LiSSA_GPT-4o_CoT_None_None_perArtifactMetrics.csv',
                 filename.itrust = 'LiSSA/iTrust_LiSSA_GPT-4o_CoT_None_None_perArtifactMetrics.csv',
                 approach = 'LiSSA', category = 'LLM', separator = ';')

tlr.lsi <- load.tlr.results(filename.etour = 'LSI/eTour_LSI_0.19_perArtifactMetrics.csv',
                 filename.itrust = 'LSI/iTrust_LSI_0.29_perArtifactMetrics.csv',
                 approach = 'LSI', category = 'Traditional', separator = ';')

tlr.vsm <- load.tlr.results(filename.etour = 'VSM/eTour_VSM_0.15_perArtifactMetrics.csv',
                 filename.itrust = 'VSM/iTrust_VSM_0.17_perArtifactMetrics.csv',
                 approach = 'VSM', category = 'Traditional', separator = ';')
```

The results from TLR do not have data for the F2 results yet, but they can easily be calculated from the available precision and recall values.

```{r calculate-f2}
calc_f_beta <- function(beta, precision, recall) {
  f <- (1+beta^2) * ((precision * recall)/(beta^2*precision + recall))
  return(ifelse(is.nan(f), 0, f))
}

tlr.ftlr <- tlr.ftlr %>% 
  mutate( F2 = calc_f_beta(2, Precision, Recall) )
```

Finally, merge the TLR results from the different approaches together.

```{r merge-tlr-results}
df.dep <- rbind(tlr.ftlr, tlr.ir, tlr.lissa, tlr.lsi, tlr.vsm)
```


The resulting data frame contains the following values.

```{r data-dependent-head}
head(df.dep)
```

## Finalization

Finally, merge the dependent and independent data frames on the `Dataset` and `File` columns. 

```{r data-merge}
df <- merge(df.ind, df.dep, by=c('Dataset', 'File'), all=TRUE)
```

Then, store the file in a location accessible to the data analysis scripts.

```{r data-storing}
write_csv(df, file=file.path('../data/rq4tlr-data.csv'))
```

