---
title: "Data Preparation"
author: "Anonymous"
date: '2025-02-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

This notebook prepares the observational data for its subsequent analysis.

## Indepedent Data

Firstly, load the `csv` sheet containing all observations, which contains both manually and automatically annotated variables.

```{r data-loading}
df.ind <- read.csv('../../data/output/rq4tlr-merged.csv')
```

### Datatype Correction

Several variables are Boolean but parsed as characters (`"TRUE"` and `"FALSE"`) in the data sheet.
To mitigate this, cast these variables to actual Boolean values.

```{r data-casting-bool}
vars.bool <- c("duplicate", "nameproblem", "scope", "textorder", "iounquantified", "actoractor", "justification", "happy", "meanuc", "lostfr")

to_bool <- function(value) {
  return(value == "True")
}

df.ind <- df.ind %>% 
  mutate_at(vars.bool, to_bool)
```

The resulting independent data contains all variables in their correct format.

```{r data-status}
str(df.ind)
```

### Value Correction

The values in the `coherence` column contain the value -1.0 whenever the use case subflow contained no more than 1 step.
In this case, the coherence value is not meaningful because there cannot be coherence of any sentence to any predecessor.

```{r coherence-hist}
df.ind %>% ggplot(aes(x = coherence)) +
  geom_histogram()
```

The value -1.0 would be misinterpreted by a regression model.
Replacing those values with `NA` will discard them for the regression analysis.
Instead, we replace those values with 1: a subflow with only one step has perfect coherence by default.

```{r coherence-adoption}
df.ind <- df.ind %>% 
  mutate(coherence = ifelse(coherence == -1, 1, coherence))
```

## Dependent Data

The dependent variables are located in the `./data/TLR_results/` directory.
The trace link recovery results per data set are stored in separate files. 

```{r data-dependent}
df.dep.etour <- read.csv('../../data/TLR_results/FTLR/etour_ORG_m59f44_FTLRMcCd_perArtifactMetrics.csv', sep=';') %>%
  mutate(Dataset = 'etour')
df.dep.itrust <- read.csv('../../data/TLR_results/FTLR/itrust_ORG_m59f44_FTLRMcCd_perArtifactMetrics.csv', sep=';') %>% 
  mutate(Dataset = 'itrust')

df.dep <- rbind(
  df.dep.itrust,
  df.dep.etour
)
```

The `sourceID` column associates each row of the results with the requirements file that the TLR algorithm attempted to trace.
However, the values in this column still contain the `.txt` file extension.
To harmonize the file IDs of the dependent and independent data points, remove the file extensions.

```{r data-dependent-file}
df.dep <- df.dep %>% 
  rowwise() %>% 
  mutate(File = strsplit(sourceID, '\\.')[[1]][1])
```

Then, drop the now unnecessary `sourceID` column.

```{r data-dependent-drop}
df.dep <- df.dep %>%
  select(!(sourceID))
```

The resulting data frame contains the following values.

```{r data-dependent-head}
head(df.dep)
```

## Finalization

Finally, merge the dependent and independent data frames on the `Dataset` and `File` columns. 

```{r data-merge}
df <- merge(df.ind, df.dep, by=c('Dataset', 'File'), all=TRUE)
```

Then, store the file in a location accessible to the data analysis scripts.

```{r data-storing}
write_csv(df, file=file.path('../data/rq4tlr-data.csv'))
```

