---
title: 'Observational Analysis: Estimate Visualization'
author: "Anonymous"
date: '2025-03-05'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

path.data <- "./../data/estimates/"
path.figures <- "./../../figures/results/coefficients/"
```

This notebook visualizes the distributions of the coefficients for each factor and each response variable metric involved in the observational studies.
The observational study currently involves the following metrics.

```{r metrics}
metrics <- c("precision", "recall", "f1", "f2")
```

## Data Preparation

First, load the respective data.
Every observational analysis produced one such file per response variable metric.

```{r load-data}
d.precision <- read.csv(file.path(path.data, "obs-precision.csv"))
d.recall <- read.csv(file.path(path.data, "obs-recall.csv"))
d.f1 <- read.csv(file.path(path.data, "obs-f1.csv"))
d.f2 <- read.csv(file.path(path.data, "obs-f2.csv"))
```

Then, define a function that calculates the 95% and 50% confidence interval from the mean and variance.

```{r calculate-cis}
calculate_ci <- function(df) {
  return(df %>% mutate(
    ci1.low = mean - 2*var,
    ci1.up = mean + 2*var,
    ci2.low = mean - var,
    ci2.up = mean + var
  ))
}

d.precision <- calculate_ci(d.precision)
d.recall <- calculate_ci(d.recall)
d.f1 <- calculate_ci(d.f1)
d.f2 <- calculate_ci(d.f2)
```

Add the response variable metric as a column in each data frame.

```{r explicate-metric}
d.precision$resp = "precision"
d.recall$resp = "recall"
d.f1$resp = "f1"
d.f2$resp = "f2"
```

Finally, combine the individual datasets to one.

```{r combine-data}
d <- rbind(d.precision, d.recall, d.f1, d.f2)
```

## Data Visualization

Visualize the coefficients in an errorbar plot.

```{r figure-configuration}
dodge = 0.7 # distance between errorbars of each metric

ci1.width = 0.6 # size of the whiskers of the outer confidence interval
ci2.width = 0.6 # size of the whiskers of the inner confidence interval

ci1.thic = 0.1 # thickness of the outer confidence interval
ci2.thic = 1.3 # thickness of the inner confidence interval
```

### Full Visualization

The first figure displays the coefficient distributions of all factors, regardless of their significance.

```{r visualize-all}
d %>% 
  ggplot(aes(x=coef, color=resp)) +
  geom_errorbar(aes(ymin = ci1.low, ymax = ci1.up), width=ci1.width, linewidth=ci1.thic, position=position_dodge(width=dodge)) +
  geom_errorbar(aes(ymin = ci2.low, ymax = ci2.up), width=ci2.width, linewidth=ci2.thic, position=position_dodge(width=dodge)) +
  coord_flip() +
  geom_hline(yintercept=0, color = "grey")
```

```{r export-full}
ggsave(
  file = file.path(path.figures, "obs-coefs-full.pdf"), device = "pdf", 
  width = 16, height = 20, units = "cm")
```

### Significant Visualization

For better understandability, also create a visualization only involving the significant factors.
A factor is considered significant if any of the metric distributions has a inner confidence interval that is not consistent with 0.

```{r filter-significant}
significant_coefs <- d %>% 
  filter(coef != "Intercept") %>% # ignore the Intercept coefficient
  mutate(significant = !(ci2.low < 0 & ci2.up > 0)) %>% # determine a coefficient as significant if the inner (50%) confidence interval does not overlap with zero
  group_by(coef) %>% # group by coefficient
  summarize(contains_significant = (sum(significant)>0)) %>% # determine all coefficients where at least one of the measured metrics is significant
  filter(contains_significant) # keep only these coefficients
```

Display the significant factors.

```{r visualize-significant}
d %>% 
  filter(coef %in% significant_coefs$coef) %>% 
  ggplot(aes(x=coef, color=resp)) +
  geom_errorbar(aes(ymin = ci1.low, ymax = ci1.up), width=ci1.width, linewidth=ci1.thic, position=position_dodge(width=dodge)) +
  geom_errorbar(aes(ymin = ci2.low, ymax = ci2.up), width=ci2.width, linewidth=ci2.thic, position=position_dodge(width=dodge)) +
  coord_flip() +
  geom_hline(yintercept=0, color = "grey")
```

```{r export-significant}
ggsave(
  file = file.path(path.figures, "obs-coefs-sig.pdf"), device = "pdf", 
  width = 16, height = 10, units = "cm")
```
