---
title: "marginal-approaches"
author: "Anonymous"
date: "2025-08-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(RColorBrewer)

path.data <- "./../data/estimates/"
path.figures <- "./../../figures/results/coefficients/"
```

This notebook visualizes the baseline performance of the five approaches regarding the four studied metrics (precision, recall, $F_1$-score, and $F_2$-score).

```{r metrics}
metrics <- c("Precision", "Recall", "F1", "F2")
approaches <- c("VSM", "LSI", "FTLR", "IR", "LiSSA")
```

## Data Preparation

First, load the respective data.
Every observational analysis produced one such file per response variable metric.

```{r load-data}
d.precision <- read.csv(file.path(path.data, "marginal-approaches-precision.csv"))
d.recall <- read.csv(file.path(path.data, "marginal-approaches-recall.csv"))
d.f1 <- read.csv(file.path(path.data, "marginal-approaches-f1.csv"))
d.f2 <- read.csv(file.path(path.data, "marginal-approaches-f2.csv"))
```

Then, define a function that calculates the 95% and 50% confidence interval from the mean and variance.

```{r calculate-cis}
calculate_ci <- function(df) {
  return(df %>% mutate(
    ci1.low = estimate__ - 2*se__,
    ci1.up = estimate__ + 2*se__,
    ci2.low = estimate__ - se__,
    ci2.up = estimate__ + se__
  ))
}

d.precision <- calculate_ci(d.precision)
d.recall <- calculate_ci(d.recall)
d.f1 <- calculate_ci(d.f1)
d.f2 <- calculate_ci(d.f2)
```

Add the response variable metric as a column in each data frame.

```{r explicate-metric}
d.precision$resp = "Precision"
d.recall$resp = "Recall"
d.f1$resp = "F1"
d.f2$resp = "F2"
```

Finally, combine the individual datasets to one and put the approach column in order

```{r combine-data}
d <- rbind(d.precision, d.recall, d.f1, d.f2)

d <- d %>% 
  mutate(Approach = factor(d$Approach, levels = approaches, ordered = TRUE)) %>% 
  mutate(Approach = fct_rev(Approach)) %>% 
  mutate(resp = factor(d$resp, levels = metrics, ordered = TRUE)) %>% 
  mutate(resp = fct_rev(resp))
```

The `forcats::fct_rev` is necessary since horizontal errorbars with a `color` aesthetic will show the factor levels in reverse order otherwise.

## Data Visualization

Visualize the coefficients in an errorbar plot.

```{r figure-configuration}
dodge = 0.6 # distance between errorbars of each metric

ci1.width = 0.6 # size of the whiskers of the outer confidence interval
ci2.width = 0.6 # size of the whiskers of the inner confidence interval

ci1.thic = 0.7 # thickness of the outer confidence interval
ci2.thic = 1.2 # thickness of the inner confidence interval
```

```{r visualize-distribution}
ggplot(d, aes(y = resp, color=Approach)) +
  geom_point(aes(x = estimate__), position=position_dodge(width=dodge), size = 2) + 
   # error bar for the "outer" confidence interval
    geom_errorbar(
      aes(xmin = ci1.low, xmax = ci1.up),
      width=ci1.width, linewidth=ci1.thic,
      position=position_dodge(width=dodge)) +
    # error bar for the "inner" confidence interval
    geom_errorbar(
      aes(xmin = ci2.low, xmax = ci2.up),
      width=ci2.width, linewidth=ci2.thic,
      position=position_dodge(width=dodge)) +
    # vertical line around 0 to show the threshold for significance
    geom_vline(xintercept=0, color = "grey") +
    theme_bw() +
    theme(axis.title.y=element_blank()) + 
    labs(x = "Average expected value") +
    scale_color_manual(values = brewer.pal(n = 5, name = "Set1"), limits = approaches)
```

```{r export-figure}
ggsave(
  file = file.path(path.figures, "obs-marginal-approaches.pdf"), device = "pdf", 
  width = 18, height = 10, units = "cm")
```
