---
title: "Marginal Effects of Tangled Requirements"
author: "Anonymous"
date: "2025-08-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(RColorBrewer)

path.data <- "./../data/estimates/"
path.figures <- "./../../figures/results/marginal/"
```

This notebook visualizes the marginal effect of *Tangled Requirements* (`outlinks`) on the four studied metrics (precision, recall, $F_1$-score, and $F_2$-score).

```{r metrics}
metrics <- c("Precision", "Recall", "F1", "F2")
```

## Data Preparation

First, load the respective data.
Every observational analysis produced one such file per response variable metric.

```{r load-data}
d.precision <- read.csv(file.path(path.data, "marginal-outlinks-precision.csv"))
d.recall <- read.csv(file.path(path.data, "marginal-outlinks-recall.csv"))
d.f1 <- read.csv(file.path(path.data, "marginal-outlinks-f1.csv"))
d.f2 <- read.csv(file.path(path.data, "marginal-outlinks-f2.csv"))
```

Finally, combine the individual datasets to one and put the approach column in order

```{r combine-data}
d <- rbind(d.precision, d.recall, d.f1, d.f2)

d <- d %>% 
  mutate(metric = factor(d$metric, levels = metrics, ordered = TRUE))# %>% 
  #mutate(metric = fct_rev(metric))
```

The `forcats::fct_rev` is necessary since horizontal errorbars with a `color` aesthetic will show the factor levels in reverse order otherwise.

## Data Visualization

Visualize the marginal effects in a line plot with ribbons.

```{r visualize-plot}
ggplot(d, aes(x = outlinks)) + 
  geom_line(aes(y = mean, color = metric)) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = metric), alpha = 0.1) +
  labs(x = "Number of outgoing Traceability Links", y = "Average Metric Value", color = "Metric", fill = "Metric") +
  scale_color_manual(values = brewer.pal(n = 4, name = "Set1"), limits = metrics) +
  scale_fill_manual(values = brewer.pal(n = 4, name = "Set1"), limits = metrics) +
  scale_x_continuous(breaks = c(0, 3, 6, 9, 12)) +
  ylim(0.18, 0.6) +
  theme_bw() +
  theme(legend.position = "bottom")
```

```{r export-figure}
ggsave(
  file = file.path(path.figures, "obs-marginal-outlinks.pdf"), device = "pdf", 
  width = 4, height = 3)
```
