---
title: "Experimental Analysis"
author: "Anonymous"
date: '2025-02-26'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

library(brms)
library(marginaleffects)
```

In this Notebook, we conduct an experimental analysis.
As opposed to the observational analysis, the experimental analysis allows to fully control a factor and, hence, isolate its effect on the response variable of interest.

```{r load-data}
df <- read.csv('./data/rq4tlr-data.csv')

head(df)
```


## Factor Selection

We can only conduct an experimental analysis of a specific factor if

1. the factor is sufficiently wide distributed in the data, and
2. we can reliably manipulate the factor without requiring domain knowledge.

Only one factor in our data meets both criteria: the ratio of sentences representing clarifications (`clarification`).
The factor is at least spread a little in the `iTrust` data set.

```{r visualize-clarifications}
df %>% 
  filter(Dataset == "itrust") %>% 
  ggplot(aes(x = clarification, fill = Dataset)) +
  geom_histogram(stat="count")
```

```{r}
df.treatment <- read.csv('../data/TLR_results/FTLR/itrustClarificationCleaned_ORG_m59f44_FTLRMcCd_perArtifactMetrics.csv', sep=';') %>%
  mutate(Dataset = 'itrust') %>% 
  rowwise() %>% 
  mutate(File = strsplit(sourceID, '\\.')[[1]][1]) %>% 
  select(!c(sourceID, tp, fp, fn))

df.treatment$clarification = 0.0

head(df.treatment)
```

```{r}
df.control <- df %>% 
  filter(Dataset == "itrust") %>% 
  select(c(Dataset, File, clarification, Precision, Recall, F1, AP, LAG))

head(df.control)
```

```{r merge-data}
d <- rbind(df.control, df.treatment)
```


## Data Analysis

```{r outcome-distribution}
d %>% 
  select(c(Precision, Recall, F1)) %>% 
  pivot_longer(cols=c(Precision, Recall, F1), names_to = "Metric") %>% 
  ggplot(aes(x = value, fill=Metric)) +
    geom_histogram()
```

```{r formula}
f <- bf(F1 ~ 0 + Intercept + clarification + (1|File))
```

```{r}
get_prior(f, d, family=zero_one_inflated_beta)
```



```{r select-priors}
p1 <- c(
  prior(normal(0, 0.5), class = b),
  prior(gamma(4, 1), class = phi),
  prior(beta(5, 2), class = coi), 
  prior(beta(3, 2), class = zoi),
  prior(weibull(2, 0.5), class = sd)
)
```

```{r model-sample-priors}
m1.prior <-
  brm(data = d, family = zero_one_inflated_beta, f, prior = p1,
    iter = 4000, warmup = 1000, chains = 4, cores = 4,
    seed = 4, sample_prior="only",
    file = "fits/exp.f1.prior"
  )
```

```{r prior-predictive-checks}
brms::pp_check(m1.prior, ndraws=100)
```


```{r model}
m1 <-
  brm(data = d, family = zero_one_inflated_beta, f, prior = p1,
    iter = 4000, warmup = 1000, chains = 4, cores = 4, seed = 4,
    file = "fits/exp.f1"
  )
```

```{r}
brms::pp_check(m1, ndraws=100)
```


```{r path-coefficients}
summary(m1)
```

```{r}
mcmc_plot(m1, variable=c("b_clarification", "sd_File__Intercept"))
```

```{r marginal}
conditional_effects(m1, effects="clarification", method="posterior_epred")
```
